---
title: 'MY472 Final Exam Part C: Scraping additional data'
---

Scraping the data available on: https://everypolitician.org/united-states-of-america/house/term-table/114.html and https://everypolitician.org/united-states-of-america/senate/term-table/114.html

```{r message=FALSE, warning=FALSE}
library(rvest)
library(reshape2)
library(tidyverse)
library(data.table)
library(eeptools) #Calculate age
library(ggplot2)

congress <- read.csv("congress-list.csv", stringsAsFactors = F)
neg_com <- read.csv("neg_com.csv", stringsAsFactors = F)

# Write html tables and bind all resulting tables to two separate dataframes
url <- "https://everypolitician.org/united-states-of-america/house/term-table/114.html"
tab_house <- html_table(read_html(url), fill = TRUE)
tab_house <- rbindlist(tab_house) # Convert to dataframe

url2 <- "https://everypolitician.org/united-states-of-america/senate/term-table/114.html"
tab_senate <- html_table(read_html(url2), fill = TRUE)
tab_senate <- rbindlist(tab_senate) # Convert to dataframe
```

```{r warning=FALSE}
# Combine both house and senate dataframes into one
leg_tab <- rbind(tab_house, tab_senate)

# Keep only necessary rows of information
leg_tab <- leg_tab %>%
  filter(X1 == "Gender" | X1 == "Born" | X1 == "bioguide")

# Cast the data from long to wide format using row ID numbers
leg_tab <- dcast(leg_tab, rowid(X1) ~ X1, value.var = "X2")

# Calculate age based on system date
leg_tab$Born <- as.Date(leg_tab$Born)

leg_tab <- leg_tab %>%
  mutate(Age = age_calc(Born, enddate = Sys.Date(),
                        units = "years", precise = F))

```
Measuring negativity on pages of younger politicians
```{r}
# Merge the legislator table with congress csv using the bioguide
neg_pol <- merge(x = leg_tab,
               y = congress,
               by.x = "bioguide", by.y = "bioguide_id")

neg_pol <- merge(x = neg_pol,
                 y = neg_com,
                 by = "facebook")

# Create a scatterplot with a linear regression
# to determine relation of age and negativity
scatter <- ggplot(neg_pol, aes(x = Age,
                               y = percent_negative, label = from_name))
fit <- lm(percent_negative ~ Age, data = neg_pol)

summary(fit)

scatter +
  scale_x_continuous() +
  scale_y_continuous() +
  geom_point(aes(alpha = abs(residuals(fit)))) +
  geom_smooth(method = lm) +
  guides(alpha = F) +
  theme_minimal()

```

**Based on the medians of gender, party, and a combination of gender and party, there is no significant difference in negative comments between and within the variables. However, within the combination of gender and party, Independent male representatives have the most negativity, while Democrat males have the least negativity.**

**Looking at variance for each group, male legislators have a larger variance of negative comments at 19.2% versus female legislators at 12.3%. Republicans have a larger variation than Democrats, with Independent legislators having the most (due to small sample size the likelihood of larger variance is higher). Combining both party and gender, male Republicans have the most variance (discounting Independent males as, once again, the population is small) and Democrat females have the least amount of variance.**

**As a result, we can say that there doesn't appear to be an obvious correlation between the number of gender, party, and negative comments received. However, Democrats, specificially Democrat women, seem to have a narrower range of percentage of negative comments received than Republicans.**

```{r}
# Because of the range of negativity in legislators, use median
neg_pol %>%
  group_by(Gender) %>%
  summarise(gender_med = median(percent_negative),
            gender_var = var(percent_negative))

neg_pol %>%
  group_by(party) %>%
  summarise(party_med = median(percent_negative),
            party_var = var(percent_negative))

neg_pol %>%
  group_by(party, Gender) %>%
  summarise(party_gen_med = median(percent_negative),
            party_gen_var = var(percent_negative))
```

